\section{Wordpress}\index{wordpress}

Wordpress is a PHP content management system. It has reasonable security in a new default install, but has a poor track record for remote execution exploits. The best way to secure Wordpress is to update it if possible and remove all unnecessary or old plugins.

\subsection{Setup}

Download the latest tarball available at \lstinline|https://wordpress.org/latest.tar.gz| and untar it into the document root (i.e. /var/www). Create and configure the necessary using the following SQL commands.

\acmlisting[caption=Wordpress SQL Setup, label=Wordpress SQL Setup]{./services/wordpress/sql/wordpress.sql}

Navigate to the setup page at \lstinline|http://localhost/wordpress/| and follow the setup instructions. Proceed below with how to add a few extra layers of security to a Wordpress installation.

\subsection{Security}

\begin{itemize}
	\item make sure file permissions are restrictive
	\begin{itemize}
		\item `/` needs to be writable only by the owning user account (e.g. 'www')
		\item `/wp-content` needs to be writable by web server (e.g. 'apache')
		\item `/wp-content/plugins` needs to be writable only by the owning user account (e.g. 'www')
	\end{itemize}

	\item remove unnecessary database permissions
	\begin{itemize}
		\item reduce database permissions for the SQL user by running the following command, replacing \lstinline|wordpress.*| with the Wordpress tables and \lstinline|wordpress@localhost| with the Wordpress user if necessary.
		\item \lstinline|REVOKE ALL PRIVILEGES ON wordpress.* from wordpress@localhost; GRANT SELECT,INSERT,UPDATE,DELETE ON wordpress.* TO wordpress@localhost; FLUSH PRIVILEGES;|
	\end{itemize}

	\item disable file editing from wp-admin
	\begin{itemize}
		\item add \lstinline|define('DISALLOW_FILE_EDIT', true);| to \lstinline|wp-config.php|
	\end{itemize}

	\item move \lstinline|wp-config.php| to the directory above the Wordpress root

	\item add 'AskApache Password Protect' which enables HTTP authentication preventing wp-admin from being exploited
\end{itemize}
