\section{ssh}\index{ssh}\index{sshd}

SSH or Secure Shell is a remote administration protocol.
It allows the user to send remote commands to Linux machines (and soon versions of windows 10 or later).
It can be a vary powerful tool for system administration, but can also be a powerful exploit target if unsecured.
It is used with a variety of tools, including the provisioning tool Ansible and back up tool rsync for secure connections.
In general there are a few best practices to follow for using ssh:

\begin{itemize}
	\item Disable root login.
	\item Disable password authentication (default).
	\item Disable host-based authentication (default).
	\item Ensure that ssh is not setuid to prevent host-based authentication.
	\item Use at least public key authentication, Use 2-factor authentication on where possible.
	\item During competition revoke all authorized keys except the scoring engine as required.
	\item Use sandbox privilege separation to prevent privilege escalation attacks on the daemon (default).
	\item Use PAM (pluggable authentication module) (default).
	\item Block excessive connections to ssh at the firewall.
	\item Do not forward the SSH Agent to untrusted/compromised servers.
\end{itemize}

Important system level configuration directories and files
\begin{itemize}
	\item /etc/sshd/sshd\_config --- daemon configuration.
	\item /etc/hosts.equiv --- used for insecure host based authentication; remove when found.
	\item /etc/shosts.equiv --- used for insecure host based authentication; remove when found.
	\item /etc/ssh/ssh\_known\_hosts --- system wide list of host keys.
	\item /etc/ssh/ssh\_host\_*key --- private keys used for host-based authentication and fingerprints.
	\item /etc/ssh/sshrc --- commands that are executed when the user logs on.
\end{itemize}

Important user level configuration directories and files
\begin{itemize}
	\item \~{}/.rhosts --- used for insecure host based authentication; remove when found.
	\item \~{}/.shosts --- used for insecure host based authentication; remove when found.
	\item \~{}/.ssh/known\_hosts --- list of hosts that are not already in /etc/ssh/ssh\_known\_hosts.
	\item \~{}/.ssh/authorized\_keys/ --- list of keys that can be used to authenticate as this user.
	\item \~{}/.ssh/config --- per user configuration options for ssh.
	\item \~{}/.ssh/environment --- environment options for the user.
	\item \~{}/.ssh/id*.pub --- public key for the user.
	\item \~{}/.ssh/id* --- private key for the user.
\end{itemize}

This script will set most of the sane settings described above for Linux Boxes:
\acmlisting[caption=Sshd Setup Script, label=Sshd Setup Script]{./services/ssh/scripts/sshd-config.sh}
